Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-01-06 grove-paper.tex
add tree syntax as a grammar
commit draft notes from last Grove meeting

@@

Finally, any remaining single-parented vertices are parts of wreaths. A wreath
is a cycle of single-parented vertices that each may have single-parented
descendants.

@@

>>>>

% XXX

(illustrative running example: 1 of each: mp, d, r, sc)

declarative section: give a hint what decomposition does (incl. wreaths): take a
graph, turn it into a grove (i.e., 4-tuple) which is what we actually present to
users and use for things like type checking. In this section, we present an
algorithm that declaratively say what a grove is and how it connects to the
graph.

We can't turn the graph into a single tree, so we turn it into a forest. A
forest is a collection of trees, in fact it's a 4-tuple consisting of a rooted tree
and 3 sets of trees: mp, d, sc. Before discussing what these sets are, let's 
define what a tree is.

here's what a tree is. why do we care what a tree is? because we can map them to
the expressions in figure 11. Explain what a Ref is wrt forest roots. (Maybe
mention that only well-sorted trees can be mapped, which we discuss below.)

There's this weird stuff left over; we call them wreaths and put them in the 4th
tuple component.

here's what a wreath is (= a connected graph where each vertex has exactly one
parent)

(show what a wreath looks like)

here's what a grove is

here are some properties that show how to get one from the other, and to/from
graph

algorithmic section: if we want to implement these conversions, here are the
steps (without introducing new concepts):

- linear scan
- breaking out roots
- ...

@@

Finally, once we've already accounted for things with more or less than one
parent, everything else must be single parented. We call these things wreaths. A
wreath is a cycle of single-parented vertices that each may have single-parented
descendants.

lemma: a wreath is a cycle with trees hanging off of it

lemma: wreath traversal always produces a valid tree

@@

% XXX

decomposition produces sets of trees. As long as everything is well sorted, we
can use them to produce expressions in the main grammar. (alt: define wrt the
graph instead, once we establish the graph can be turned into expressions)

*** well-sortedness judgment on trees: the tree only has valid positions for the
constructors, and the child at each position has the sort corresponding to its
position, and edges in position maps all start at the given constructor.

``A graph has this property iff these conditions hold''

then we can define a total mapping from well-sorted trees to expressions
(because the tree structure is very generic, so this ties them back to the
earlier grammar.)

forest = 4-tuple

forest validity: each of the trees in the forest are well sorted, and all Refs
point to roots of other trees of the correct sort.

every valid forest corresponds to a well behaved grove (i.e., ``expression
forest'')

soundness: if you start with a valid graph, you get a well behaved grove that
connects every vertex (and only those vertices) in the graph. (i.e., if you turn
it back into a graph, it's the same graph (modulo edge IDs unless we add $\e$ to
Refs??? OR make Ref nullary and pull edge/vertex info from the position map and
add Root sort

OR:

use edge IDs instead of edges and build a mapping from vertices to
sorts---change the $\e$s to $u$s)

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Trees:\qquad}  & t & {}\in Tree & {}::={} & \textrm{Con}(v, m) \mid \textrm{Ref}(v) \\
      \textrm{Maps:}         & m & {}\in Map & {}::={} & [\,] \mid (p \mapsto [(\e, t), \ldots, (\e, t)]); m \\
    \end{array}
  \]
  \caption{Syntax of graph decomposition trees as a grammar.}
  \label{fig:Syntax of graph decomposition trees}
\end{figure}
